---
title: "TS exercise on VARs SVARs"
author: "HR"
date: "11/16/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Define a hypothesis:

## Load the data:

```{r}
library(mFilter)
library(urca)
library(dynlm)
library(strucchange)
library(vars)
library(readxl)
library(readr)
data <- read.csv("dataset.csv")
```

## Plot the data:

```{r}
hp1 <- ts(data$house.prices, start=c(1950,1), frequency = 4)
y1 <- ts(data$real.gdp, start=c(1950,1), frequency = 4)
r <- ts(data$national.bank.lending.rate , start=c(1950,1), frequency = 4)
c1 <- ts(data$credit, start=c(1950,1), frequency = 4)

hp <- log(hp1)
y <- log(y1)
c <- log(c1)

plot(hp)
plot(y)
plot(r)
plot(c)
```

## Test for seasonality

```{r}
X <- cycle(y)
q1 <- ifelse(X == "1", 1, 0)
q2 <- ifelse(X == "2", 1, 0)
q3 <- ifelse(X == "3", 1, 0)
q4 <- ifelse(X == "4", 1, 0)
summary(y.reg <- lm(y ~ q1 + q2 + q3))
```

## Deasonalised vs seasonal data

We did not find any evidence of seasonal fluctuation

## Unit root testing

```{r}
arpdiag <- function(series, lags) {
x <- mat.or.vec(lags, 1)
y <- mat.or.vec(lags, 1)
y <- y + 0.05
for (i in 1:lags) {
x[i] <- Box.test(series, lag = i, type = "Ljung-Box")$p.value
}
plot(x, xlab = "Lags", ylab = "p-value H0: no
Autocorrelation", type = "p",
main = "Ljung-Box Test for
Autocorrelation", ylim = c(0, 1))
axis(1, 1:lags)
abline(0.05, 0, lty = 2, col = "blue")
}
```

### House prices

**general-to-specific metode**

*Start*

```{r}
model1_start<-ur.df(hp,lags=12,type='trend')
summary(model1_start)
```

*End*
```{r}
model1<-ur.df(hp,lags=9,type='drift')
res<- model1@testreg$residuals
arpdiag(res, 12)
summary(model1)
```


**Omvendt, anbefaldes IKKE (af Enders i hvertfald, se side 217)**

*Start*

```{r}
model1_start<-ur.df(hp,lags=6,type='drift')
summary(model1_start)
```

*End*
```{r}
model1<-ur.df(hp,lags=6,type='drift')
res<- model1@testreg$residuals
arpdiag(res, 12)
summary(model1)
```

**Alternative, keep significant lags:**

*Start*

```{r}
library(dynlm)
model_start_hp <- dynlm(diff(hp) ~ 1+ L(hp, 1) + L(diff(hp), c(1:12))+ trend(diff(hp)))
summary(model_start_hp)
```

*test*

```{r}
library(stringr)
library(car)
vars_hp <- str_c("L(diff(hp), c(1:12))", c(10:12, 2:3, 5, 7 ))
linearHypothesis(model_start_hp, vars_hp, rep(0, length(vars_hp)))
```

*End*

```{r}
model_end_hp <- dynlm(diff(hp) ~ 1+ L(hp, 1) + L(diff(hp), c(1,4,6,8,9)))
summary(model_end_hp)
res= residuals(model_end_hp)
arpdiag(res,12)

summary(ur.df(hp,lags=6,type='drift'))
```





### Real GDP

**Start**
```{r}
model2<-ur.df(y,lags=12,type='trend')
summary(model2)
```


**End**
```{r}
model2<-ur.df(y,lags=8,type='trend')
res<- model2@testreg$residuals
arpdiag(res, 12)
summary(model2)
model2@teststat
```

### Credit

**Start**

```{r}
model3<-ur.df(c,lags=12,type='trend')
summary(model3)
```


**End**
```{r}
model3<-ur.df(c,lags=8,type='drift')
res<- model3@testreg$residuals
arpdiag(res, 12)
summary(model3)
model2@teststat
```

### Interest rate

**Start**

```{r}
model4<-ur.df(r,lags=2,type='none')
summary(model4)
```

**End**
```{r}
model4<-ur.df(r,lags=2,type='none')
res<- model4@testreg$residuals
arpdiag(res, 12)
summary(model4)
model2@teststat
```

### Make the series stationary

```{r}
y_d <- diff(y)
hp_d <- diff(hp)
r_d <- diff(r)
c_d <- diff(c)
```

```{r}
model5<-ur.df(y_d,lags=8,type='trend')
res<- model5@testreg$residuals
arpdiag(res, 12)
summary(model5)
model5@teststat
model5@cval
```

```{r}
plot(y_d)
```


```{r}
model6<-ur.df(hp_d,lags=8,type='trend')
res<- model6@testreg$residuals
arpdiag(res, 12)
summary(model6)
model6@teststat
model6@cval
```

```{r}
plot(hp_d)
```


```{r}
model7<-ur.df(c_d,lags=4,type='trend')
res<- model7@testreg$residuals
arpdiag(res, 12)
summary(model7)
model7@teststat
model7@cval
```

```{r}
plot(c_d)
```

```{r}
model8<-ur.df(r_d,lags=4,type='trend')
res<- model8@testreg$residuals
arpdiag(res, 12)
summary(model8)
model8@teststat
model8@cval
```

```{r}
plot(r_d)
```

All my 4 variables are I(1)

```{r}
dataset <- cbind(y_d, r_d, c_d, hp_d)
VARselect(dataset)
```


## Granger causality

```{r}
library(car)
yg <- dynlm(y_d~L(y_d, c(1:5)) + L(r_d, c(1:5)) + L(c_d, c(1:5)) + L(hp_d, c(1:5) ))
summary(yg)
yg1 <- dynlm(y_d~L(y_d, c(1:5)) + L(c_d, c(1:5)) + L(hp_d, c(1:5) ))
anova(yg, yg1)
yg2 <- dynlm(y_d~L(y_d, c(1:5)) + L(r_d, c(1:5))  + L(hp_d, c(1:5) ))
anova(yg, yg2)
yg3 <- dynlm(y_d~L(y_d, c(1:5)) + L(r_d, c(1:5))  + L(c_d, c(1:5) ))
anova(yg, yg3)
```


## VAR model

```{r}
var <- VAR(y = dataset, p = 5, type = 'const')
roots(var)
serial.test(var, lags.pt=12)
arch.test(var, lags.multi = 12, multivariate.only=T)
normality.test(var)
```

```{r, out.width=1}
plot(stability(var, type="OLS-CUSUM"))
```


## Estimate SVAR

```{r}
bmat<-diag(4) 
diag(bmat)<-1 
bmat[2,1]<-NA 
bmat[3,1]<-NA 
bmat[3,2]<-NA
bmat[4,1]<-NA 
bmat[4,2]<-NA 
bmat[4,3]<-NA
bmat
```


```{r}
svar <- SVAR(x = var, Bmat = bmat)
```

## Impulse responses

```{r}
shock1 = irf(svar, n.ahead=24, impulse='y_d', ci=0.66, response=c('r_d'), cumulative=T)
shock2 = irf(svar, n.ahead=24, impulse='y_d', ci=0.66, response=c('c_d'), cumulative=T)
shock3 = irf(svar, n.ahead=24, impulse='y_d', ci=0.66, response=c('hp_d'), cumulative=T)
shock4 = irf(svar, n.ahead=24, impulse='hp_d', ci=0.66, response=c('y_d'), cumulative=T)
```


```{r}
par(mfrow=c(3,3))
plot(shock1, xlab="") 
plot(shock2, xlab="") 
plot(shock3, xlab="")
plot(shock4, xlab="")
```

## Is the effect of house price shocks on GDP asymmetrics (non-linear)?

```{r}
pos_diff <- function(x) {
     xd <- x - diff(x)
     y<- cumsum((0 + 1*(xd > 0))*xd)
}

neg_diff <- function(x) {
  xd <- x - diff(x)
  y<- cumsum((0 + 1*(xd < 0))*xd)
}

hp_pos <- pos_diff(hp_d)
hp_neg <- neg_diff(hp_d)
hp_pos <- ts(hp_pos, start=c(1950,1), frequency = 4)
hp_neg <- ts(hp_neg, start=c(1950,1), frequency = 4)
```
Plot negative and positive movements separately
```{r}
plot(hp_pos)
plot(hp_neg)
hp_pos_d<- diff(hp_pos)
hp_neg_d <- diff(hp_neg)
plot(hp_pos_d)
plot(hp_neg_d)
```


```{r}
dataset <- cbind(y_d, r_d, c_d, hp_pos_d)
dataset <- na.omit(dataset)
VARselect(dataset)
```

```{r}
var1 <- VAR(y = dataset, p = 5, type = 'const')
roots(var1)
svar1 <- SVAR(x = var1, Bmat = bmat)
```

## Effect of positive shocks to house prices on GDP

```{r}
shock1 = irf(svar1, n.ahead=24, impulse='hp_pos_d', ci=0.66, response=c('y_d'), cumulative=T)
plot(shock1)
```


```{r}
dataset <- cbind(y_d, r_d, c_d, hp_neg_d)
dataset <- na.omit(dataset)
VARselect(dataset)
```

```{r}
var1 <- VAR(y = dataset, p = 5, type = 'const')
roots(var1)
svar1 <- SVAR(x = var1, Bmat = bmat)
```

## Effect of negative shocks to house prices on GDP

```{r}
shock1 = irf(svar1, n.ahead=24, impulse='hp_neg_d', ci=0.66, response=c('y_d'), cumulative=T)
plot(shock1)
```
