---
title: "Lecture 8"
author: "Simon"
date: "3/11/2022"
output: html_document
---

# Spurious regressions 

```{r}
set.seed(2113)
n = 100
e = replicate(4, {
rnorm(100, 0, 1)
})
e1 = e[, 1]
e2 = e[, 2]
e3 = e[, 3]
e4 = e[, 4]
Y = c(1)
X1 = c(1)
X2 = c(1)
X3 = c(1)
t = 1:n
for (t in 2:n) {
Y[t] <- Y[t - 1] + e1[t]
X1[t] <- X1[t - 1] + e2[t]
X2[t] <- X2[t - 1] + e3[t]
X3[t] <- X3[t - 1] + e4[t]
}
```

```{r}
par(mfrow = c(2, 2))
plot(Y, t = "l")
plot(X1, t = "l")
plot(X2, t = "l")
plot(X3, t = "l")

```

```{r}
library(dynlm)
summary(dynlm(Y ~ X1 + X2 + X3))
```

```{r}
summary(dynlm(Y ~ L(Y, 1) + X1 + X2 + X3))
```

```{r}
set.seed(2113)
n=80
e= replicate(4, {rnorm(100,0,1)} )
e1=e[,1]; e2=e[,2]
mu=0.5
Y=c(1); X=c(1);
t=1:n
for (t in 2:n) {
Y[t] <- mu + Y[t-1] + e1[t]
X[t] <- mu + X[t-1] + e2[t]
}


```

```{r}
summary(dynlm(Y ~ X))
```

```{r}
summary(dynlm(Y ~ X + L(Y,1)))
```


# Cointegration



```{r}
n = 100
set.seed(2113)
e = replicate(4, {
rnorm(100, 0, 1)
})
e1 = e[, 1]
e2 = e[, 2]
Y = c(1)
X = c(1)
t = 1:n
for (t in 2:n) {
X[t] <- X[t - 1] + e1[t]
Y[t] <- 0.8 * X[t] + e2[t]
}
```



```{r}
PP.test(Y)
PP.test(X)

```

```{r}
PP.test(diff(Y))
PP.test(diff(X))
```

**Johansen cointegration**

```{r}
model <- lm(Y ~ X)
error = residuals(model)

```

```{r}
arpdiag <- function(series, lags) {
x <- mat.or.vec(lags, 1)
y <- mat.or.vec(lags, 1)
y <- y + 0.05
for (i in 1:lags) {
x[i] <- Box.test(series, lag = i, type = "Ljung-Box")$p.value
}
plot(x, xlab = "Lags", ylab = "p-value H0: no
Autocorrelation", type = "p",
main = "Ljung-Box Test for
Autocorrelation", ylim = c(0, 1))
axis(1, 1:lags)
abline(0.05, 0, lty = 2, col = "blue")
}
```


```{r}
library(urca)
model1<-ur.df(error,lags=0,type='none')
res<- model1@testreg$residuals
arpdiag(res, 12)

```

```{r}
model1<-ur.df(error,lags=0,type='none')
model1@testreg

```

## Exercise 

Get data on two variables that you believe should have a long-run relationship
I Saving and Investment, Disposable income and consumption, real
exchange rate and exports, etc. are a few examples in macro showing
long-run relationship
Test the two variables for a unit root.
Test for cointegration.


# Error correction model 

```{r}
Y = ts(Y)
X = ts(X)
summary(dynlm(diff(Y) ~ L(diff(Y), 1) + L(diff(X), 1) + diff(X) + L(Y,
1) + L(X, 1)))
```

```{r}
dynlm(diff(Y)~diff(X) + L(Y,1) + L(X,1))
```



```{r}
Y = ts(Y)
X = ts(X)
model = dynlm(Y ~ X)
model 

error = residuals(model)
```


```{r}
ecm= dynlm(diff(Y)~diff(X) + L(error,1))
summary(ecm)

```

## Excercise 

