---
title: "Lecture 2"
output: html_document
date: "2022-09-20"
---

# calculating cyclical correlation 


```{r}
set.seed(524)
a <- rnorm(80); x <- cumsum(a)
x <- ts(x, start = c(2010, 1), frequency = 4)


set.seed(525)
b <- rnorm(80); y <- cumsum(b);
y <- ts(y, start = c(2010, 1), frequency = 4)
```


```{r}
plot(x)
plot(y)
```




```{r}
x <- as.numeric(x) # this converts the variable from ts() to a normal vector (the ccf function below works 
y <- as.numeric(y)
cc <- ccf(x, y, 1, pl = F)
cc
```



## Cyclical volatility

```{r}
vy <- (hpy/hptrend) * 100
sd(vy)
```

```{r}
ccf(y, x, 1, pl = F)
```

```{r}
cc=ccf(x, y, 2, pl=F)
plot(cc, ci=0.95, ylab="", xlab="", main="") # ci=0.95 means 95 percent confidence interval.
```


## Calculating cyclical volatility 

```{r}
set.seed(525)
a <- rnorm(80, mean = 1.5, sd=3)
x<- cumsum(a)
plot(x, t="l")
```


We created a trend, we can remove this and use the cyclical volatility 

```{r}
library(mFilter)
lambda <- 1600
hp = hpfilter(x, lambda)
trend_x <- hp$trend
```



```{r}
v_x <- ((x - trend_x) * 100/trend_x)
```


```{r}
sd(v_x)
```




# Exercise 

1. Which variables are procyclical and which ones are counter-cyclical?

2. Which business cycle variable is the most volatile?
-Note: Remember to work with seaonsally adjusted and detrended data!


Step 1.  Load data

```{r}
library(readxl)
data <- read_excel("data.xlsx")
data=lapply(data, function(t) ts(t, start=c(1950, 1), end=c(2018, 1),frequency=4))
attach(data)

```


**step 2. deseasonalyze the data**


```{r}
X <- cycle(real_Credit)
q1 <- ifelse(X == "1", 1, 0)
q2 <- ifelse(X == "2", 1, 0)
q3 <- ifelse(X == "3", 1, 0)
q4 <- ifelse(X == "4", 1, 0)


summary(real_Credit.reg <- lm(real_Credit ~ q1 + q2 + q3)); mean_real_cred= mean(real_Credit)
summary(stock_prices.reg <- lm(`stock prices` ~ q1 + q2 + q3)); mean_stock_price= mean(`stock prices`)
summary(real_house_prices.reg <- lm(`real house prices` ~ q1 + q2 + q3)); mean_real_house= mean(`real house prices`)
summary(real_gdp.reg <- lm(real_gdp ~ q1 + q2 + q3)); mean_real_gdp= mean(real_gdp)

```


```{r}
real_cred_de=ts(residuals(real_Credit.reg)+mean_real_cred, start = 1950, frequency = 4)
stock_price_de=ts(residuals(stock_prices.reg)+ mean_stock_price, start = 1950, frequency = 4)
real_house_de=ts(residuals(real_house_prices.reg)+mean_real_house , start = 1950, frequency = 4)
real_gdp_de=ts(residuals(real_gdp.reg)+mean_real_gdp , start = 1950, frequency = 4)
```

```{r}
plot(real_cred_de); lines(real_Credit, col= "red")
plot(stock_price_de); lines(`stock prices`, col= "red")
plot(real_house_de); lines(`real house prices`, col= "red")
plot(real_gdp_de); lines(real_gdp, col= "red")
```



**Step 3 detrend**


```{r}
library(mFilter)

# REal credit

lambda <- 1600
hp1 = hpfilter(real_cred_de, lambda)
hptrend1 <- hp1$trend


#Stock prices
hp2 = hpfilter(stock_price_de, lambda)
hptrend2 <- hp2$trend

#Real house
hp3 = hpfilter(real_house_de, lambda)
hptrend3 <- hp3$trend

#real gdp
hp4 = hpfilter(real_gdp_de, lambda)
hptrend4 <- hp4$trend
```


```{r}

# REal credit
real_cred_det = ts((real_Credit - hptrend1), start=1950, frequency=4)
real_cred_det1 <- real_cred_det + real_Credit[1]

#Stock prices
stock_price_det = ts((`stock prices` - hptrend2), start=1950, frequency=4)
stock_price_det1 <- stock_price_det + `stock prices`[1]

#Real house

real_house_det = ts((`real house prices` - hptrend3), start=1950, frequency=4)
real_house_det1 <- real_house_det + `real house prices`[1]
#real gdp

real_gdp_det = ts((real_gdp - hptrend4), start=1950, frequency=4)
real_gdp_det1 <- real_gdp_det + real_gdp[1]
```


```{r}
ts.plot(real_Credit, hptrend1, real_cred_det1, lwd=c(1,1,1), lty=c(3,1,1), xlab="Year", col=c("black", "red", "blue"))

ts.plot(`stock prices`, hptrend2, stock_price_det1, lwd=c(1,1,1), lty=c(3,1,1), xlab="Year", col=c("black", "red", "blue"))

ts.plot(`real house prices`, hptrend3, real_house_det1, lwd=c(1,1,1), lty=c(3,1,1), xlab="Year", col=c("black", "red", "blue"))

ts.plot(real_gdp, hptrend4, real_gdp_det1, lwd=c(1,1,1), lty=c(3,1,1), xlab="Year", col=c("black", "red", "blue"))
```



**Step 4 look at cross correlation**

```{r}
#Real credit

cc1=ccf(as.numeric(real_cred_det1), as.numeric(real_gdp_det1), 2, pl=F)
plot(cc1, ci=0.95, ylab="", xlab="", main="")
```



