---
title: "Lecture 4"
author: "Simon"
date: "2/10/2022"
output: html_document
---

# ARIMA(p,q) process


# Test for stationarity

## DF-test

```{r}
set.seed(123123)
Yt <- arima.sim(list(order = c(0, 1, 0)), n = 100)
TT <- 100
Yt <- ts(cumsum(rnorm(TT)))

```


```{r}
library(dynlm)
summary(dfuller.reg <- dynlm(diff(Yt) ~ 0 + L(Yt, 1))) # 0 means no intercept in the regression
```



```{r}
library(Quandl)
bnp = Quandl("FRED/GDPC1", start_date="1970-01-01", end_date="2018-12-01", type="ts")
bnp=ts(bnp, start=c(1970,1), end=c(2018,4), frequency = 4) # format as a time series object


plot(bnp)
```


```{r}
Ytd=diff(Yt) # We take the first difference of Yt
```

```{r}
summary(dfuller.reg <- dynlm(diff(Ytd) ~0 + L(Ytd, 1)))

```


### Exampple Testing US GDP

```{r}
# install.packages("Quandl")
library(Quandl)
bnp = Quandl("FRED/GDPC1", start_date="1970-01-01", end_date="2018-12-01", type="ts")
bnp=ts(bnp, start=c(1970,1), end=c(2018,4), frequency = 4) # format as a time series object
plot(bnp)
```

**DF test with no constant or intercept**

```{r}
library(dynlm)
summary(dfuller.reg <- dynlm(diff(bnp) ~ 0 + L(bnp, 1))) # 0 means no intercept in the regression
```
**DF test with a constant**
```{r}
library(dynlm)
summary(dfuller.reg <- dynlm(diff(bnp) ~ 1 + L(bnp, 1))) # 0 means no intercept in the regression
```

```{r}
library(dynlm)
summary(dfuller.reg <- dynlm(diff(bnp) ~ 1 + L(bnp, 1) + trend(diff(bnp))))

```



We fail to reject that pi=0 and that the series has unit root. 
So we take the first difference 


```{r}
Ytd=diff(bnp) # We take the first difference of US GDP
```

```{r}
library(dynlm)
summary(dfuller.reg <- dynlm(diff(Ytd) ~0 + L(Ytd, 1)))

```
```{r}
summary(dfuller.reg <- dynlm(diff(Ytd) ~1 + L(Ytd, 1)))
```

```{r}
summary(dfuller.reg <- dynlm(diff(Ytd) ~1 + L(Ytd, 1) + trend(Ytd)))

```


### Exercise

Get data for quarterly GDP for Denmark, and perform a unit root test
using DF test.





```{r}
library(readxl)
Danish_GDP <- read_excel("Danish_GDP.xlsx")

bnp_dk= ts(Danish_GDP$bnp, start= c(1995,1), end = c(2022,2), frequency = 4)
```



First seasonalize the data: 

```{r}
X <- cycle(bnp_dk)
q1 <- ifelse(X == "1", 1, 0)
q2 <- ifelse(X == "2", 1, 0)
q3 <- ifelse(X == "3", 1, 0)
q4 <- ifelse(X == "4", 1, 0)
```


```{r}
summary(bnp_dk.reg <- lm(bnp_dk ~ q1 + q2 + q3)); mean_bnp_dk= mean(bnp_dk)
```

```{r}
bnp_dk_de=ts(residuals(bnp_dk.reg)+mean_bnp_dk, start =c(1995,1), frequency = 4)
```



```{r}
summary(dfuller.reg <- dynlm(diff(bnp_dk_de) ~ 0 + L(bnp_dk_de, 1)))
```


We add a constant: 

```{r}
summary(dfuller.reg <- dynlm(diff(bnp_dk_de) ~ 1 + L(bnp_dk_de, 1)))
```


We add a trend 
```{r}
summary(dfuller.reg <- dynlm(diff(bnp_dk_de) ~ 1 + L(bnp_dk_de, 1)+ trend(bnp_dk_de)))

plot(bnp_dk_de)
```


Looks like the variable is trend stationary!



## ADF test

```{r}
library(dynlm)
library(urca)
library(forecast)
library(car)
library(stringr)
```


```{r}
model<-dynlm(diff(bnp)~1 + L(bnp,1) + trend(diff(bnp)) )
res=residuals(model)
acf(res); pacf(res)

```

We can see the series include serial correlation 



```{r}
model_start <- dynlm(diff(bnp) ~ 1 + L(bnp, 1) + L(diff(bnp), 1:12) + trend(diff(bnp)))
summary(model_start)

```

You can remove lags 1 by 1 as in slides 
Or use the F-test to remove more lags,
BUT you need to test for normality and serial correlation to make sure the standard error is not biased!
We will test for serial correlation in the end. 


```{r}
res_start= residuals(model_start)
shapiro.test(res_start)
```


Seems like we should fix normality, but im lazy atm. 

F-test

```{r}
 vars_1 <- str_c("L(diff(bnp), 1:12)", c(10:12, 3:7))
```

```{r}


vars_out_1 <- c(vars_1)
vars_out_1
 
linearHypothesis(model_start, vars_out_1, rep(0, length(vars_out_1)))

```


```{r}
model <- dynlm(diff(bnp) ~ 1 + L(bnp, 1) + L(diff(bnp), c(1,2,8,9)) + trend(diff(bnp)))
summary(model)
```



We can now test for serial correlation, we use the funcktion Hamid created 

```{r}
arpdiag <- function(series, lags) {
x <- mat.or.vec(lags, 1)
y <- mat.or.vec(lags, 1)
y <- y + 0.05
for (i in 1:lags) {
x[i] <- Box.test(series, lag = i, type = "Ljung-Box")$p.value
}
plot(x, xlab = "Lags", ylab = "p-value H0: no
Autocorrelation", type = "p",
main = "Ljung-Box Test for
Autocorrelation", ylim = c(0, 1))
axis(1, 1:lags)
abline(0.05, 0, lty = 2, col = "blue")
}
```


```{r}
res_model= residuals(model)
arpdiag(res_model, 12)
```
We have removed serial correlaion, lets go!




### Exercise 


**Perform Augmented-DF (ADF) test on Danish GDP Make sure there is no serial correlation in your model when you compare your t statistics values with the critical values in Enders Make sure you increase the power of the test by including as less number of lags as possible (you can either start with minimum no. of lags and keep adding lags OR you can start with the highest number of lags and start reducing step by step.)**

```{r}
model_start <- dynlm(diff(bnp_dk_de) ~ 1 + L(bnp_dk_de, 1) + L(diff(bnp_dk_de), 1:12) + trend(diff(bnp_dk_de)))
summary(model_start)
```

```{r}
 vars_1 <- str_c("L(diff(bnp_dk_de), 1:12)", c(2:12))
```

```{r}


vars_out_1 <- c(vars_1)
vars_out_1
 
linearHypothesis(model_start, vars_out_1, rep(0, length(vars_out_1)))

```


```{r}
model <- dynlm(diff(bnp_dk_de) ~ 1 + L(bnp_dk_de, 1) + L(diff(bnp_dk_de), c(1)) + trend(diff(bnp_dk_de)))
summary(model)

library(urca)

model <- ur.df(bnp_dk_de, lags = 1, type = "trend")
summary(model)
```
We need to test for serial correlation to make sure the standard error is not biased. 

```{r}
res_model= model@testreg$residuals
arpdiag(res_model, 12)
```

Seems like one lag was enough! But we fail to reject H0 that there is unit root, and will now take the first difference. 



```{r}
diff_bnp_dk_de= diff(bnp_dk_de)
```

```{r}
model_start <- dynlm(diff(diff_bnp_dk_de) ~ 1 + L(diff_bnp_dk_de, 1) + L(diff(diff_bnp_dk_de), 1:12))
summary(model_start)
```

```{r}
 vars_1 <- str_c("L(diff(diff_bnp_dk_de), 1:12)", c(1:12))
```

```{r}


vars_out_1 <- c(vars_1)
vars_out_1
 
linearHypothesis(model_start, vars_out_1, rep(0, length(vars_out_1)))

```


```{r}
model <- dynlm(diff(diff_bnp_dk_de) ~ 1 + L(diff_bnp_dk_de, 1))
summary(model)
```

```{r}
res_model= residuals(model)
arpdiag(res_model, 12)
```
No serial correlation, so it seems like its an I(1) variable!



**What is you conclusion regarding unit root?**

The varialbe is I(1) as we have to differentiate to make it stationary




**Which stochastic process best describes the Danish GDP, randomwalk with drift or a pure random walk? What is your educated guess before even performing the test?**

Random walk with a drift, As we can see the mean is increasing over time!
And we can see the mean of the differentiated variable is different than 0

```{r}
plot(bnp_dk_de)

plot(diff_bnp_dk_de);mean(diff_bnp_dk_de)
```



## Phillips Peron test

```{r}
TT <- 200
x = rnorm(TT)
Yt <- ts(cumsum(x))
PP.test(Yt)

```


```{r}
plot(bnp);abline(v=2009, lwd=2, col="red")
```

## Structural breaks

```{r}
library(urca)
test <- ur.za(bnp, model = c("both"), lag=12)
summary(test)
```


## Exercise 

**Perform PP test on Danish GDP**

```{r}
PP.test(bnp_dk_de)
```


**Perform unit root structural break on Danish GDP**

```{r}
struc_model <- ur.za(bnp_dk_de, model = c("both"), lag=1)
summary(struc_model)
```

```{r}
res_struc_model= struc_model@testreg$residuals
arpdiag(res_struc_model, 12)
```


We need to differentiate

```{r}
struc_model_diff <- ur.za(diff_bnp_dk_de, model = c("intercept"), lag=0)
summary(struc_model_diff)
```

```{r}
res_struc_model_diff= struc_model_diff@testreg$residuals
arpdiag(res_struc_model_diff, 12)
```

**Is there a break in trend?**


**Is there a break in intercept?**

**Is there a unit root in the data?**

Yes!