---
title: "Lecture 5"
output: html_document
date: "2022-10-08"
---


# The general ARMA(p,q) process


# The box Jenkins Methodology

```{r}
library(forecast)
library(strucchange)
library(Quandl)
library(tstools)
```


```{r}
bnp = Quandl("FRED/GDPC1", start_date="1970-01-01", end_date="2018-12-01", type="ts")
bnp=ts(bnp, start=c(1970,1), end=c(2018,4), frequency = 4) # format as a time series object
plot(bnp)
```


```{r}
bnp=ts(bnp, start=c(1970,1), end=c(2018,4), frequency = 4) ; Ytd=diff(bnp)
acf(Ytd); pacf(Ytd)
```



Hamid guesses its ARIMA(2,1,1)

```{r}
arima212 = Arima(bnp, order = c(2, 1, 2))
AIC(arima212)
BIC(arima212)
```

```{r}
arima211 = Arima(bnp, order = c(2, 1, 1))
AIC(arima211)
BIC(arima211)
```
```{r}
arima210 = Arima(bnp, order = c(2, 1, 0))
AIC(arima210)
BIC(arima210)
res210 = residuals(arima210)

```

```{r}
auto.arima(bnp, trace = TRUE)
```


```{r}
mymodel = auto.arima(bnp)
mymodel
```

Stationarity:

```{r}
autoplot(mymodel)
```



## Exercise

**Use the Danish quartely GDP that you have worked with in the previous lectures**

```{r}
library(readxl)
Danish_GDP <- read_excel("Danish_GDP.xlsx")

bnp_dk= ts(Danish_GDP$bnp, start= c(1995,1), end = c(2022,2), frequency = 4)
```

First seasonalize the data: 

```{r}
X <- cycle(bnp_dk)
q1 <- ifelse(X == "1", 1, 0)
q2 <- ifelse(X == "2", 1, 0)
q3 <- ifelse(X == "3", 1, 0)
q4 <- ifelse(X == "4", 1, 0)
```


```{r}
summary(bnp_dk.reg <- lm(bnp_dk ~ q1 + q2 + q3)); mean_bnp_dk= mean(bnp_dk)
```

```{r}
bnp_dk_de=ts(residuals(bnp_dk.reg)+mean_bnp_dk, start =c(1995,1), frequency = 4)
```





**Which ARIMA model best describes your time series?**

```{r}
minmodel = auto.arima(bnp_dk_de)
minmodel
```


```{r}
autoplot(minmodel)
```




### Diagnostics

**Serial correlation**


```{r}
library(stats)
```

```{r}
par(mar=c(2,2,3,1)); par(mfrow=c(1,3))
tsdiag(mymodel, gof.lag=12) # tests the model up to 12 lags for serial correlation
```


```{r}
res=residuals(mymodel)
Box.test(res, lag = 20, type = "Ljung-Box")
```


**ARCH**

```{r}
arpdiag <- function(series, lags) {
x <- mat.or.vec(lags, 1)
y <- mat.or.vec(lags, 1)
y <- y + 0.05
for (i in 1:lags) {
x[i] <- Box.test(series, lag = i, type = "Ljung-Box")$p.value
}
plot(x, xlab = "Lags", ylab = "p-value H0: no
Autocorrelation", type = "p",
main = "Ljung-Box Test for
Autocorrelation", ylim = c(0, 1))
axis(1, 1:lags)
abline(0.05, 0, lty = 2, col = "blue")
}

```


```{r}
arpdiag((res^2), lags=12)

```


**Normality**


```{r}
shapiro.test(res)

```


```{r}
checkresiduals(res)
```

Ser ud til det er 2008 Q4


Create dummy


```{r}
d_2008=create_dummy_ts(start_basic = c(1970, 1), end_basic=c(2018,4),
dummy_start=c(2008,4), dummy_end = c(2008,4),
basic_value = 0, dummy_value = 1, frequency = 4)


mymodel1 = Arima(bnp, order = c(2, 1, 0), seasonal = list(order = c(1, 0, 2)),
include.drift=TRUE, xreg=d_2008) # xreg is used to include exogenous regressors


checkresiduals(minmodel)
```



```{r}
library(tseries)

shapiro.test(residuals(mymodel1))
jarque.bera.test(residuals(mymodel1))

```


**Strucural break check**

```{r}
break_point <- breakpoints(Ytd ~ 1)
summary(break_point)
```

```{r}
plot(Ytd); lines(fitted(break_point, breaks = 1), col = 4)
lines(confint(break_point, breaks = 1))
```



# Arima forecast


Forecasting model with 2008 dummy

```{r}
ahead <- 20 # future values for forecasting
futurVal <- forecast(mymodel1, h=ahead, level=c(97.5), xreg = rep(0, ahead))
plot(futurVal)

```


Forecasting without 2008 dummy
```{r}
ahead <- 24 # future values for forecasting
pred <- forecast(mymodel, h=ahead)
plot(pred, xlim=c(2000,2020), ylim=c(12000, 20000))

```



```{r}
hold <- window(bnp, start=c(2016,1))
bnp_a <- window(bnp, end=c(2015,4))
mymodel_a= Arima(bnp_a, order = c(2, 1, 0), seasonal = list(order = c(1, 0, 2)), include.drift=TRUE)
pred <- forecast(mymodel_a, h=12)
plot(pred, xlim=c(2000,2018), ylim=c(12000, 20000), lwd=3)
lines(bnp, col="red", lwd=3)

```





