---
title: "Lecture 3"
author: "Simon"
date: "27/9/2022"
output: html_document
---

# Lecture 2

## Autoregresive process

```{r}
set.seed(213)
n=500
e <- rnorm(n, mean = 0, sd = 1.5)
Y <- rnorm(1)
for (t in 2:n) {
Y[t] <- 0.1*Y[t-1]+e[t]
}
ts.plot(Y, main=expression(paste("AR(1) process with ", theta, "=0.1")), xlab="")
```

```{r}
ar.sim <- arima.sim(model = list(ar = c(0.6, -0.3)), n = 100) + 5
plot(ar.sim)
```


# Lecture 3

##Quick recap

```{r}
TT <- 500; mean=0; sd=1
Xt <- ts(rnorm(TT, mean, sd), start=1, freq=1)
Ct <- ts(cumsum(rnorm(TT)), start=1, freq=1)
ts.plot(Ct, Xt, lty=2:1)
legend("topleft", c("Ct is non stationary", "Xt is stationary"), lty=2:1)
```



## Random walk

**Without drift or trend**

```{r}
set.seed(154)
e = rnorm(500,0,1) # e is white noise
y <- 0
for (t in 2:500) {
y[t] <- y[t-1] + e[t]
}
yd = ts(diff(y), start=1, frequency = 1) # I make it stationay by differncing
ts.plot(y[-1], yd, lty=2:1, main = "Random walk process without drift", xlab="")
```


**With drift**

```{r}
set.seed(154)
e = rnorm(500,0,1) # e is white noise
y <- 0
mu <- 0.1 # stochastic trend is + so the series will trend upwards
for (t in 2:500) {
y[t] <- mu + y[t-1] + e[t]
}
yd = ts(diff(y), start=1, frequency = 1) # I make it stationay by differncing
ts.plot(y[-1], yd, main = "Random walk process with drift", xlab="")
```



**With trend**


```{r}
set.seed(213)
n=500
e <- rnorm(n, mean = 0, sd = 1.5)
Y <- rnorm(1)
mu <- 0.1
t=1:n
for (t in 2:n) {
Y[t] <- mu + 0.5*Y[t-1] + e[t] + 0.1*t
}

yd = ts(diff(Y), start=1, frequency = 1) # I make it stationay by differncing

ts.plot(Y[-1], yd, main = "Random walk process with drift", xlab="")
```



## Lag operators

Exercise: Can you invert an MA(1) process into AR(1)?

- Start by writting the MA(1) model: 

$$Y_t= \varepsilon_t - aL \varepsilon_t$$

We can move $\varepsilon_t$ outside the parenthsis:

$$Y_t= \varepsilon_t (1-aL)$$
We can now isolate $\varepsilon_t$

$$\frac{Y_t}{(1-aL)}= \varepsilon_t$$

We can now use the geometric rule: 


$$\varepsilon_t= (aL)^0*Y_t + (aL)^1*Y_t + (aL)^2*Y_t + (aL)^n*Y_t$$

We can apply the rule for Lag operator: 

$$\varepsilon_t= Y_t + a^1*Y_{t-1} + a^2*Y_{t-2} + a^n*Y_{t-n}$$
Isolate $Y_t$

$$Y_t= \varepsilon_t - a^1* Y_{t-1} - a^2 * Y_{t-2} - a^n * Y_{t-n}$$




## Autocorrelation function (ACF)


```{r}
TT <- 200
acf(arima.sim(model=list(ar=0.9), n=TT), ylab=expression(Y[t]),
main=expression(paste("AR model with ", theta, "=0.9")))
```

```{r}
TT <- 200
acf(arima.sim(model=list(ar=-0.9), n=TT), ylab=expression(Y[t]),
main=expression(paste("AR model with ", theta, "=-0.9")))
```



```{r}
TT <- 200
acf(arima.sim(model=list(ma=0.9), n=TT), ylab=expression(Y[t]),
main=expression(paste("MA model with ", theta, "=0.9")))
```


```{r}
TT <- 200
acf(arima.sim(model=list(ma=-0.9), n=TT), ylab=expression(Y[t]),
main=expression(paste("MA model with ", theta, "=-0.9")))
```




## Partial correlation function (PACF)


```{r}
TT <- 200
x= arima.sim(model=list(ar=0.9), n=TT)
acf(x, ylab=expression(Y[t]), main= expression(paste("ACF of AR model with ", theta, "=0.9")))
pacf(x, ylab=expression(Y[t]), main= expression(paste("PACF of AR model with ", theta, "=0.9")))
```


```{r}
TT <- 200
x= arima.sim(model=list(ar=-0.9), n=TT)
acf(x, ylab=expression(Y[t]), main= expression(paste("ACF of AR model with ", theta, "=-0.9")))
pacf(x, ylab=expression(Y[t]), main= expression(paste("PACF of AR model with ", theta, "=-0.9")))
```





```{r}
TT <- 200
x= arima.sim(model=list(ma=0.9), n=TT)
acf(x, ylab=expression(Y[t]), main= expression(paste("ACF of MA model with ", theta, "=0.9")))
pacf(x, ylab=expression(Y[t]), main= expression(paste("PACF of MA model with ", theta, "=0.9")))
```


```{r}
TT <- 200
x= arima.sim(model=list(ma=-0.9), n=TT)
acf(x, ylab=expression(Y[t]), main= expression(paste("ACF of MA model with ", theta, "=-0.9")))
pacf(x, ylab=expression(Y[t]), main= expression(paste("PACF of MA model with ", theta, "=-0.9")))
```


## Exercise 

Generate a white noise series, an AR(1) series and an AR(2) series. Compare the ACFs and PACFs of these 3 series. Do the ACFs and PACFs actually help you to distinguish between the different series? That is, are you able to identify the order of the AR(p) processes by looking at their ACFs and PACFs?



**White noise** 

```{r}
set.seed(123)

e <- rnorm(n, mean = 0, sd = 1.5)
pacf(e)
acf(e)
```


**AR(1)**


```{r}
ar_1= arima.sim(model=list(ar=0.5), n=200)
pacf(ar_1)
acf(ar_1)
```



**AR(2)**

```{r}
ar_2= arima.sim(model=list(ar=c(0.5,0.4)), n=200)
pacf(ar_2)
acf(ar_2)
```



Use the ACF and PACF to describe two actual data series. Take two of the series that you used in Exercise #1.


```{r}
library(readxl)
data <- read_excel("data.xlsx")
data=lapply(data, function(t) ts(t, start=c(1950, 1), end=c(2018, 1),frequency=4))

credit= real_Credit
house_p= `real house prices`
stock_p= `stock prices`
bnp= real_gdp
```


```{r}
library(mFilter)

# credit

lambda <- 1600
hp1 = hpfilter(credit, lambda)
hptrend1 <- hp1$trend


#Stock prices
hp2 = hpfilter(stock_p, lambda)
hptrend2 <- hp2$trend

#Real house
hp3 = hpfilter(house_p, lambda)
hptrend3 <- hp3$trend

#real gdp
hp4 = hpfilter(bnp, lambda)
hptrend4 <- hp4$trend
```


```{r}

# REal credit
real_cred_det=credit - hptrend1
real_cred_det1 <- real_cred_det + credit[1]


#Stock prices
stock_price_det = stock_p- hptrend2
stock_price_det1 <- stock_price_det + stock_p[1]

#Real house

real_house_det = house_p- hptrend3
real_house_det1 <- real_house_det + `real house prices`[1]
#real gdp

real_gdp_det = bnp- hptrend4
real_gdp_det1 <- real_gdp_det + real_gdp[1]
```




**ACF and PACF**


```{r}
## Credit

acf(real_cred_det1)
pacf(real_cred_det1)
```

Ligner en AR(2)


```{r}
## Stock prices

acf(stock_price_det1)
pacf(stock_price_det1)
```


Ligner AR(2)

```{r}
## House P

acf(real_house_det1)
pacf(real_house_det1)
```

Ligner også en AR(2)

```{r}
## Bnp

acf(real_gdp_det1)
pacf(real_gdp_det1)
```