---
title: "Lektion 1"
output: html_document
date: "2022-09-13"
---


```{r}
library(stats)
library(tseries)
library(timeSeries)
```



```{r}
library(pwt)
data("pwt6.3")
td <- pwt6.3
td$rgdp <- td$rgdpch*td$pop # real GDP = real GDP per capital X population
td$hours <- td$rgdpl/td$rgdpl2th
td <- subset(td, subset=isocode=="DNK",
select=c(isocode, country, year, rgdp, rgdpch, hours, pop))
write.table(td, file="dataset1.txt")
```

```{r}
Z <- ts(td[-c(1:3)], start = 1950, frequency = 1)
```


```{r}
plot(Z[, "rgdp"], col = "black", lwd = "2", main = "Real GDP")
plot(Z[, "rgdpch"], col = "green", lwd = "2", main = "Real GDP per capita")
plot(Z[, "hours"], col = "blue", lwd = "2", main = "Hours")
plot(Z[, "pop"], col = "red", lwd = "2", main = "Population")

```





```{r}
td$productivity <- td$rgdp/td$hours
```

```{r}
DK.data <- subset(td, subset = !is.na(productivity))
DK.data <- ts(DK.data[-c(1:3)], start = min(DK.data$year), frequency = 1)
```



```{r}
library(readxl)
td <- read_excel("Danish_GDP.xlsx")
head(td)

```


```{r}
y <- ts(td$bnp, start = 1995, frequency = 4)
plot(y, xlab = "Year", ylab = "GDP", lwd = "2")

```



# Seasonality Detrending

```{r}
X <- cycle(y)
q1 <- ifelse(X == "1", 1, 0)
q2 <- ifelse(X == "2", 1, 0)
q3 <- ifelse(X == "3", 1, 0)
q4 <- ifelse(X == "4", 1, 0)
summary(y.reg <- lm(y ~ q1 + q2 + q3))

```


```{r}
plot(ts(residuals(y.reg), start = 1995, frequency = 4), lwd = 3)

```

```{r}
yhat <- residuals(y.reg) + mean(fitted.values(y.reg))
ts.plot(y, yhat, lty = c(3, 1), lwd = c(1, 1))
legend(1995, max(y), c("GDP (y)", "Seasonally adjusted GDP (yhat)"), lty = c(3,
1))

```
We can also use the last 4 quarters moving average of GDP


```{r}
library(zoo)
library(xts)
y_ma = rollmeanr(y, k = 4, fill = NA, align = "right")
plot(y_ma)
```

## Exercise

1. Take consumption (or any other) data from statistics Denmark
(seasonally unadjusted) - test for seasonality in the data




2. Adjust the data for seasonality using the dummy variable method



# Long run detrending

```{r}
bnp = ts(yhat, start = 1995, frequency = 4) # note I use seasonally adjusted series for GDP
tid = ts(1:length(bnp), start = 1995, frequency = 4)
```

```{r}
summary(y.reg <- lm(bnp ~ tid))
```



```{r}
ty = ts(residuals(y.reg),start = 1995, frequency = 4)
ty1 = ty + y[1]

plot(ty1, type= "l")
```

```{r}
trend = fitted.values(y.reg)

```


```{r}
ts.plot(bnp, trend, ty1, lty = c(1, 2), lwd = c(2, 2, 2), xlab = "Year",
col = c("black", "red", "blue"));legend(1995, max(bnp), c("Real GDP (original series)", "trend", "Real GDP (detrended series)"),lty = 1:3)
```
## Exercise

1. Now use your seasonality adjusted consumption data and test for
linear trend in the data? If there is a trend, de-trend you consumption


2. Before you analyse the series for trend, do you see trend in the data?



## HP filter


```{r}
library(mFilter)
lambda <- 1600
hp = hpfilter(bnp, lambda)
hptrend <- hp$trend
```


```{r}
hpy = ts((bnp - hptrend), start=1995, frequency=4)
hpy1 <- hpy + y[1]
```


```{r}
ts.plot(bnp, hptrend, hpy1, lwd=c(1,1,1), lty=c(3,1,1), xlab="Year", col=c("black", "red", "blue"))
legend(1995, max(bnp), c("GDP (y)", "HP-trend", "Detrended GDP (hpy)"),
lty=c(3,1,1), col=c("black", "red", "blue"))

```


## differentiating 


```{r}
dy=diff(bnp, 1)
ts.plot(hpy,dy,ty, lwd=c(1,1,1), col=c('red','black', 'blue') ,xlab="year",main="Cyclical Output")
```

### Exercise

1. Now de-trend your consumption via differencing and plot the three
detrended time series

2. Now take the Danish GDP data and calculate output gap?

3. There was a financial crisis in 2008, some researchers use the
pre-crisis trend in GDP to calculate output loss as a result of the crisis
Assume you are employed by IMF and you are given the task to
calculate the output loss of Denmark as a result of the financial crisis
in 2008. Can you plot and report the output loss of Denmark?





### Exercise

```{r}
bnp <-ts(bnp, start =c(1995, 1), freq = 4)

plot(bnp)
```


Can you find the impact of Covid-crisis on Danish real GDP?

```{r}
bnp_pre =window(bnp, start =c(2010, 1), end =c(2019, 4), freq = 4)
```



```{r}
lambda = 1600
hp =hpfilter(bnp_pre, lambda)
trend = hp$trend
```


```{r}
length(trend)
```


```{r}
mean(diff(log(trend)))
```

```{r}
n =c(1:10)
x = 535.1649 * (1 + 0.00466)^n
```

```{r}
bnp1 <-ts(bnp, start=c(1995,1), end=c(2022,2), freq=4)
bnp1[101:110] = x
```

```{r}
plot(bnp1, col="red", t="l",  lwd=3);lines(bnp, t="l", lwd=3)
```

